---
title: "varioscan analysis"
author: "Jurre Hageman"
date: '2020-02-21'
version: '2021-02-16'
output:
  html_document: default
  word_document: default
---

## User Settings

```{r}
# Set organism
organism <- "My Organism"

# Set file and folder
filename <- "2020-01-30.csv"
folder <- "sample_data"

# Set samples
sample_1 <- "Sample1"
sample_2 <- "Sample2"
sample_3 <- "Elution"

# Set columns (sample#_replicate).
# This is for user input convenience. 
# Sample and replicate will be separated later using dplyr::separate
col_1 <- "1_1" 
col_2 <- "1_2" 
col_3 <- "1_3" 
col_4 <- "1_c" 
col_5 <- "2_1" 
col_6 <- "2_2" 
col_7 <- "2_3" 
col_8 <- "2_c" 
col_9 <- "3_1" 
col_10 <- "3_2" 
col_11 <- "3_3" 
col_12 <- "3_c" 

# Set dilutions
row_1 <- 2
row_2 <- 1
row_3 <- 0.5
row_4 <- 0.25
row_5 <- 0.13
row_6 <- 0.06
row_7 <- 0.03
row_8 <- 0

# Set line colors
linecolors <- c("#FF3633", "#FF8933", "#FCFF33", "#62FF33", "#33FFEB", "#33A7FF", "#5F33FF", "#EA33FF")

```

## Imports

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Rmisc)
```

## Converting user input

```{r}
samples <- c(sample_1, sample_2, sample_3)
columns <- c("row", col_1, col_2, col_3, col_4, col_5, col_6, col_7, col_8, col_9, col_10, col_11, col_12)
dilution_series <- factor(c(row_1, row_2, row_3, row_4, row_5, row_6, row_7, row_8))
filepath <- file.path(folder, filename)
group1 <- c('1_1', '1_2', '1_3', '1_c')
group2 <- c('2_1', '2_2', '2_3', '2_c')
group3 <- c('3_1', '3_2', '3_3', '3_c')
```

## Loading the data

```{r message=FALSE, warning=False}
df <- read_csv2(filepath, 
                col_types = cols(row = 'c', .default = col_double()),
                col_names = columns) %>% #Using col_names also removes the empty columns
  slice(-1:-2) #drop the first two rows

skip <- 14 #each time, 14 rows of plate info that will be skipped
collect <- 8 #there are 8 rows of OD data that I will keep
reps <- nrow(df) %/% (skip + collect) #number of repetitions
index <- rep(0:(reps - 1), each = collect) * (skip + collect) + (1 : collect) + skip 
df <- df[index, ] #collect the blocks of OD data
```

## Add timepoint and dillution columns

```{r}
num_of_timepoints <- nrow(df) / 8 - 1
timepoints_actual <- seq(0, (10 * num_of_timepoints), by = 10)
timepoints <- rep(timepoints_actual, each = 8)
dilutions <- rep(dilution_series, times = num_of_timepoints + 1)

df <- mutate(df, timepoint = timepoints, dilution = dilutions) %>%
  select(1, timepoint, dilution, everything())
#View(df)
```

## Make tidy

```{r}
#Use dplyr to make tidy df tables
#Need to gather at multiple levels, hence the function.
generateTidy <- function(df_t, selector){
  df_t <- df_t %>% 
    select(row, timepoint, dilution, all_of(selector)) %>%
    gather(selector[1], selector[2], selector[3], key = 'sample', value = 'OD_S') %>%
    separate(sample, into = c('sample', 'replicate')) %>%
    select(1:3, sample, replicate, 'OD_S', selector[4]) %>%
    dplyr::rename('OD_C' = selector[4]) %>%
    mutate(OD_N = OD_S - OD_C)
  return(df_t)
}


df_tidy_s1 <- generateTidy(df, group1)
df_tidy_s2 <- generateTidy(df, group2)
df_tidy_s3 <- generateTidy(df, group3)

```

## Summary Data

```{r}
#Use the summarySE package to calculate the mean, SD, SE and CI data at once.
vars <- c("dilution", "timepoint")
stat.sample1 <- summarySE(df_tidy_s1, measurevar="OD_N", groupvars=vars)
stat.sample2 <- summarySE(df_tidy_s2, measurevar="OD_N", groupvars=vars)
stat.sample3 <- summarySE(df_tidy_s3, measurevar="OD_N", groupvars=vars)
```

## Set-up plots

```{r, echo=F}

base.string <- sprintf("(%s) file: %s",organism, filename)
title.string.s1 <- sprintf("%s %s", samples[1], base.string)
title.string.s2 <- sprintf("%s %s", samples[2], base.string)
title.string.s3 <- sprintf("%s %s", samples[3], base.string)

generatePlot <- function(sample, title.string){
  ggplot(data = sample, mapping = aes(x = timepoint, y = OD_N, group=dilution)) + 
  geom_point(aes(color = dilution)) + 
  geom_line(aes(color = dilution)) +
  scale_color_manual(values = linecolors) +
  labs(x = "Time (min)", y = "OD", title = title.string) +
  geom_errorbar(aes(ymin = OD_N - sd, ymax = OD_N + sd), width = .05, colour = "#A9A9A9") +
  theme(plot.title = element_text(size=12))
}

```

## Generate Plots

### Plot 1

```{r, echo=F}
generatePlot(stat.sample1, title.string.s1)
```

### Plot 2

```{r, echo=F}
generatePlot(stat.sample2, title.string.s2)
```

### Plot 3

```{r, echo=F}
generatePlot(stat.sample3, title.string.s3)
```

## Controls

```{r}
generatePlotControl <- function(sample, title.string){
  ggplot(data=sample, mapping = aes(x = timepoint, y = OD_C, group = dilution)) + 
  geom_point(aes(color = dilution)) + 
  geom_line(aes(color = dilution)) +
  scale_color_manual(values = linecolors) +
  labs(x = "Time (min)", y = "OD", title = paste("Control:", title.string, sep = " ")) +
    ylim(0, 0.5) +
    theme(plot.title = element_text(size=12))
}
```

### Plot 1

```{r, echo=F}

generatePlotControl(df_tidy_s1, title.string.s1)

```

### Plot 2

```{r, echo=F}

generatePlotControl(df_tidy_s2, title.string.s2)

```

### Plot 3

```{r, echo=F}

generatePlotControl(df_tidy_s3, title.string.s3)

```

## Losse samples

```{r}
generatePlotEachLine <- function(sample, title.string){
  ggplot(data=sample, mapping = aes(x = timepoint, y = OD_N, group = dilution)) + 
  geom_point(aes(color = dilution)) + 
  scale_color_manual(values = linecolors) +
  labs(x = "Time (min)", y = "OD", title = paste("Each sample separate:", title.string, sep = " ")) +
    theme(plot.title = element_text(size=12))
}
```

### Plot 1

```{r, echo=F}

generatePlotEachLine(df_tidy_s1, title.string.s1)

```

### Plot 2

```{r, echo=F}

generatePlotEachLine(df_tidy_s2, title.string.s2)

```

### Plot3

```{r, echo=F}

generatePlotEachLine(df_tidy_s3, title.string.s3)

```

### Growthcurver


#### data transformat
ion
```{r}
#filter(stat.sample1, dilution == "0.03") # check
d <-  stat.sample1[, c(1, 2, 4)] %>% spread(key = "dilution", value = "OD_N")
colnames(d)[1] <- "time"
d$time <- d$time / 60
d
```

```{r}
library(growthcurver)
gc_fit <- SummarizeGrowth(d$time, d$`0`, bg_correct = "none")
gc_fit
```


```{r}

plot(gc_fit, ylab = "OD", xlab = "time (hours)", main = "concentration = 0")
```
The gcfit object returned from SummarizeGrowth also contains further metrics 
summarizing the growth curve data.


```{r}
str(gc_fit$vals)
```

